Title

### 2023-10-19 main.tex
Improvements

% line numbers
\usepackage[right]{lineno}

@@

To represent the algorithms implemented in quantum computers, either to
simulate a physical system or for some other purpose, one often uses a quantum
circuit~\cite{chuangbook}.  In this work, we shall also work with parametrized
quantum circuits, that is,  quantum circuits in which some of the operations
depend on variable parameters~\cite{cerezo}.  These circuits play an important
part in applications such as quantum machine learning~\cite{Benedetti} and
describing general quantum transformations dependent on parameters. Substantial
research has been devoted to enhancing the efficiency of these
circuits~\cite{Rasmussen}.  \\

We start by providing the definition of quantum channels, the general framework
used here, and multi-qubit Pauli channels in section \ref{sec: Pauli Channels}.
Our first objective is to present a quantum algorithm capable of simulating
Pauli channels on quantum computers; we do this in section \ref{sec: Circuit
for a Pauli Channel}, where we also demonstrate its implementation using IBM's
quantum computers for the particular case of single-qubit Pauli channels. 
Expanding beyond discrete Pauli channels, we introduce the concept of Pauli
dynamical maps, defined as a continuous parametrization of multi-qubit Pauli
channels.  Therefore, in section \ref{sec: 1PR Circuits} we shift our focus to
study parametrized quantum channels, aiming to adapt the algorithm developed
for channels to dynamical maps.  Furthermore, we contribute to the body of work
related to parametrized quantum circuits by establishing a theorem,
which  sets the mathematical conditions for the transformations
that can be done using a parametrized circuit with the restriction that only a
controlled single-qubit
rotation in the circuit may depend on the parameter.  Finally, in
section \ref{sec: 1PR circuit for a Pauli map}, we conclude analyzing the Pauli
dynamical maps that fulfill the conditions of theorem \ref{theorem2}.

\revnote{I fail to see the point here - the logic of the paper is: % {{{
1."open quantum systems can be modeled through a class of channels representing
he various types of decoherence. 2. Quantum circuits are a convenient
gate-based formalism that is well understood and has an extensive body of work,
and experimentally accessible.  3. We map (generalised) Pauli channels onto
quantum circuits, as a way to simulate them. \\
What I am missing is what do we learn by doing it which can't be learned by
actually performing the Pauli channel onto a quantum system? For example,
applying well controlled sx, sy, sz operations and their coherent parametric
combination using a photonic polarisation or path encoding is experimentally
trivial. Why would I need to decompose (not even effectively for any channel)
the Pauli channels to match a gate based description? Which class of problems
can we simulate that cannot be simulated otherwise? Is there any advantage (for
example in speed up, resources like ancillas required, resistance to unwanted
noise etc..?)\\
Intuitively, it seems incredibly harder resource wise to create the sum
$b|gamma\rangle$ superposition state needed for the ancilla, just for the sake
of having it into a "quantum circuit" perspective.} % }}}
\tbnote{Entiendo que su punto es que los canales de Pauli se podrían hacer de una forma % {{{
similar pero en lugar de usar los qubits de ancilla para crear las probabilidades, crearlas
de otra forma que sea mejor y que entonces no tiene mucho sentido el artículo.  \\
La verdad no sé bien cómo contestar a este comentario. \\
Una forma podria ser cambiar un poco el enfoque. 
Hacer que el objetivo principal sea
simular un mapa dinámico (en lugar de simular canales de Pauli asi nada mas)  y
estudiar como entra el parámetro en juego. 
Luego, decir que para simular mapas parametrizados de Pauli,
necesitamos empezar con un algortimo para canales y que funcione para cualquier canal de Pauli.
Al ser tan general, no es perfecto, pues varios canales se podrian hacer más fácil, 
pero lo bueno es que es un mismo algorigmo para todos, lo que 
nos va a permitir usarlo para canales parametrizados, ya que entonces sólo habrá que
mover un parámetro, en vez de cambiar todo el algoritmo para pasar de un canal a otro.\\
Luego, generalizamos el circuito para ahora sí hacerlo en mapas dinámicos y vemos qué onda con lo de 
poder reducir el número de parámetros, con lo que
llegamos al teorema de los circuitos 1PR y etc.
También corremos el circuito en computadoras cuánticas,
lo que ayuda a ver que haber modificado el circuito para que tenga un sólo parámetro
(hacer lo de meter las matrices A y B), no afectó mucho la fidelidad.
Observar cómo se afecta o no la fidelidad al hacer eso puede ser una motivación de
haber usado compus cuanticas. \\
Al final, también recalcar que el resutlado sobre 1PR es más
general que solo canales de Pauli y puede ser útil
para cualquier curcuito parametrizado.} % }}}

Ahora sí la nueva propuesta de la intro (los primeros dos párrafos son iguales): \\
$\;$ \\
{\color{green}

To represent the algorithms implemented in quantum computers, either to
simulate a physical system or for some other purpose, one often uses a quantum
circuit~\cite{chuangbook}.

However, in this article, our main goal is not to implement Pauli channels, but to study and implement Pauli dynamical maps, which are continuous parametrized curves of Pauli channels. 
These maps can be used to model continuous change of a qubit instead of only discrete jumps.
Implementing these maps will lead us to study  quantum algorithms with free parameters, 
something that is common in areas such as quantum machine learning~\cite{Benedetti}.
This can be done using parametrized quantum circuits, which consist of quantum circuits in which one or more operations depend on a free parameter. \\

Our study of such circuits for the use of simulating Pauli dynamical maps will lead 
us to theorem 2, in which we find general condition for any parametrized
transformation (for example, a Pauli dynamical map) to be achievable on a quantum circuit 
with only one single-qubit rotation dependent on the parameter.
Said parameter is in some sense a knob that we can control to continuously perform the transformation by
only having to move one number. \\

We start the article by providing the definition of quantum channels, the general framework
used here, and multi-qubit Pauli channels in section \ref{sec: Pauli Channels}.
Then, since our main objective is to obtain a parametrized circuit for Pauli dynamical maps, we will first need a 
circuit that can implement any Pauli channel by only changing some parameters
in the quantum gates and without changing the design of the circuit itself.
Therefore, in section\ref{sec: Circuit
for a Pauli Channel} we present such a circuit
and we also show its implementation using IBM's quantum computers for the particular case of single-qubit Pauli channels and some two-qubit ones.
In section \ref{sec: 1PR Circuits}, we adapt the algorithm to dynamical maps.
Then we aim to simplify the algorithm so that it only contains one parametrized single-qubit rotation, 
and in the process we contribute to the work in parametrized quantum circuits by establishing   in
\ref{theorem2} which general quantum transformations
can be implemented this way.
Finally, in
section \ref{sec: 1PR circuit for a Pauli map}, we conclude analyzing the Pauli
dynamical maps that fulfill the conditions of theorem \ref{theorem2}. 
}

@@

We can also see how an arbitrary Pauli channel acts on an arbitrary  
density matrix.
To do it, we substitute \eref{eq: Density Matrix} 
in \eref{eq: Pauli channel 1 qbit}:
\begin{equation}
\mathcal{E}(\rho) = \dfrac{1}{2}\sum_{\gamma,\alpha=0}^3 k_{\gamma} r_{\alpha} \sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma}.
\end{equation}
This can be simplified by using the following property of Pauli matrices:
\revnote{Really confusing notation, I suggest use the much more well known relation using the Levi-Civita tensor.}

\tbnote{Siguiendo la relaci\'on de conmutaci\'on, resulta que las Levi Civitas desaparecen y quedan s\'olo deltas de Dirac:}
{\color{green}
\begin{equation*}
\sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma}  =   A_{\alpha,\gamma} \sigma_{\alpha}, \;\;\;\;\;\; \text{with} \; A_{\alpha, \gamma} = (2\delta_{ij}-1)(2\delta_{i0}-1)(2\delta_{j0}-1)
\end{equation*}
}

\begin{equation}
\label{eq: propiedad-pauli}
\sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma} = A_{\alpha,\gamma} \sigma_{\alpha}, \;\;\;\;\; \text{with} \; A = \begin{pmatrix}
1 & 1 & 1 & 1\\
1 & 1 & -1 &-1 \\
1 & -1 & 1 & -1 \\
1 & -1 & -1 & 1
\end{pmatrix},
\end{equation}
which leads to
\begin{equation}
\label{eq: rho-transformada}
\mathcal{E}(\rho) = \dfrac{1}{2} \sum_{\alpha} \left(\sum_{\gamma} A_{\alpha, \gamma} k_{\gamma} \right) r_{\alpha} \sigma_{\alpha}.
\end{equation}
\eref{eq: rho-transformada} once again has the form of \eref{eq: Density Matrix} \revnote{eq 2} \tbnote{No, sí es la eq 1}
but with components $\left( \sum_{\gamma} A_{\alpha,\gamma} k_{\gamma} \right) r_{\alpha}$.
This gives us another way of understanding Pauli channels
as operations that take each component $r_{\alpha}$
of the density matrix and multiplies them
by $\sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}$, 
that is:
\begin{equation}
\label{eq: multipliers}
r_{\alpha}  \xrightarrow[\text{Channel}]{\text{Pauli}}  \tau_{\alpha} r_{\alpha} ,\;\; \tau_{\alpha} := \sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}.
\end{equation}
Notice that $\tau_0 = 1$, which is a consequence of $\sum_{\gamma}k_{\gamma}=1$
and ensures that after the channel, the resulting density matrix still has trace one. 
Furthermore, reverting the definition of $\tau_{\alpha}$ by using that $A^{-1} = \dfrac{1}{4} A$, 
we get that $k_{\gamma} = \dfrac{1}{4} \sum_{\alpha} A_{\alpha,\gamma} \tau_{\alpha}$.
Then, using that $k_{\gamma} \geq 0$
we get the following conditions on the multipliers $\tau_{\alpha}$:
\begin{eqnarray}
\label{eq: conditions-tetrahedron}
&1+\tau_i -\tau_j - \tau_k \geq 0,  \;\;\; \text{for $i,j,k$ different numbers in $\{1,2,3\}$}, \\
&1+\tau_1 + \tau_2 + \tau_3 \geq 0.
\end{eqnarray}

@@

In this section we propose a quantum circuit that simulates $N$-qubit Pauli
channels. \revnote{At this point I am missing a proper definition of a "quantum circuit" consistent with the mathematical formalism. The authors start from the very basic regarding Pauli channels, but give for granted the quantum circuit formalism.} \tbnote{Agregué algo abajo:}\\

{\color{green} A quantum circuit is a model for representing computation on a quantum system. 
This computation may include preparing the initial state of the system, applying unitary operations and measuring the whole system or parts of it~\cite{Rieffel}.

Our systems always consists of some number of qubits,
that can be divided into the ones to which we will apply the quantum channel,
 or ``principal qubits'' and the ones that help us complete said operation, or ``ancilla qubits''.
The circuit implements a unitary operator $U$ on the 
system consisting of all the qubits 
and then the principal qubits are measured.

We will create a general circuit that can implement any $N$-qubit Pauli channel
by only changing some parameters and not changing the design of the circuit itself.
This is done with the intent of converting it into a parametrized circuit for dynamical maps in the next section.
}

 Moreover, we implement the circuit for $N=1$ on a quantum computer
and analyze the results using the diamond norm.
We find that close to the depolarizing channel, the general circuit simulator 
can implement such channels with the highest fidelity.  \\

@@

{\color{green}
Finally, we move to the main goal of this article, which is 
simulating Pauli dynamical maps.
This is pretty much already solved, since
just as Pauli channels, Pauli dynamical maps can be implemented using the circuit
of \fref{Fig1}.} However, there is one difference: 
the state to be created on the ancilla qubits now 
depends on a parameter $p$, and it is represented by the expression:
\begin{eqnarray}
\label{eq: parametrized state}
\sum_{\vec{\gamma}} b_{\vec{\gamma}}(p) |\vec{\gamma}\rangle.
\end{eqnarray}
Thus, we temporarily shift our focus from Pauli
channels and dynamical maps to the general problem of 
creating a circuit to generate a curve of
states like the one described in \eref{eq: parametrized state}.

>>>>

% line numbers
\usepackage[mathlines]{lineno}  \linenumbers \setlength\linenumbersep{3pt}

@@

To represent the algorithms implemented in quantum computers, either to
simulate a physical system or for some other purpose, one often uses a quantum
circuit~\cite{chuangbook}.  In this work, we shall also work with parametrized
quantum circuits, that is,  quantum circuits in which some of the operations
depend on variable parameters~\cite{cerezo}.  These circuits play an important
part in applications such as quantum machine learning~\cite{Benedetti} and
describing general quantum transformations dependent on parameters. Substantial
research has been devoted to enhancing the efficiency of these
circuits~\cite{Rasmussen}.

\cpnote{Iteara de nuevo Tomas:
However, in this article, our main goal is not to implement Pauli channels, but to study and implement Pauli dynamical maps, which are continuous parametrized curves of Pauli channels. 
These maps can be used to model continuous change of a qubit instead of only discrete jumps.
Implementing these maps will lead us to study  quantum algorithms with free parameters, 
something that is common in areas such as quantum machine learning~\cite{Benedetti}.
This can be done using parametrized quantum circuits, which consist of quantum
circuits in which one or more operations depend on a free parameter. }

\cpnote{Porfa mete esto en el parrafo donde va la estructura:
Our study of such circuits for the use of simulating Pauli dynamical maps will
lead us to theorem 2, in which we find general condition for any parametrized
transformation (for example, a Pauli dynamical map) to be achievable on a
quantum circuit with only one single-qubit rotation dependent on the parameter.
Said parameter is in some sense a knob that we can control to continuously
perform the transformation by only having to move one number. }

\cpnote{dejar solo un parrafo de estructura}
We start by providing the definition of quantum channels, the general framework
used here, and multi-qubit Pauli channels in section \ref{sec: Pauli Channels}.
Our first objective is to present a quantum algorithm capable of simulating
Pauli channels on quantum computers; we do this in section \ref{sec: Circuit
for a Pauli Channel}, where we also demonstrate its implementation using IBM's
quantum computers for the particular case of single-qubit Pauli channels.
Expanding beyond discrete Pauli channels, we introduce the concept of Pauli
dynamical maps, defined as a continuous parametrization of multi-qubit Pauli
channels.  Therefore, in section \ref{sec: 1PR Circuits} we shift our focus to
study parametrized quantum channels, aiming to adapt the algorithm developed
for channels to dynamical maps.  Furthermore, we contribute to the body of work
related to parametrized quantum circuits by establishing a theorem, which  sets
the mathematical conditions for the transformations that can be done using a
parametrized circuit with the restriction that only a controlled single-qubit
rotation in the circuit may depend on the parameter.  Finally, in section
\ref{sec: 1PR circuit for a Pauli map}, we conclude analyzing the Pauli
dynamical maps that fulfill the conditions of theorem \ref{theorem2}.

We start the article by providing the definition of quantum channels, the
general framework used here, and multi-qubit Pauli channels in section
\ref{sec: Pauli Channels}.  Then, since our main objective is to obtain a
parametrized circuit for Pauli dynamical maps, we will first need a circuit
that can implement any Pauli channel by only changing some parameters in the
quantum gates and without changing the design of the circuit itself.
Therefore, in section\ref{sec: Circuit for a Pauli Channel} we present such a
circuit and we also show its implementation using IBM's quantum computers for
the particular case of single-qubit Pauli channels and some two-qubit ones.  In
section \ref{sec: 1PR Circuits}, we adapt the algorithm to dynamical maps.
Then we aim to simplify the algorithm so that it only contains one parametrized
single-qubit rotation, and in the process we contribute to the work in
parametrized quantum circuits by establishing   in \ref{theorem2} which general
quantum transformations can be implemented this way.  Finally, in section
\ref{sec: 1PR circuit for a Pauli map}, we conclude analyzing the Pauli
dynamical maps that fulfill the conditions of theorem \ref{theorem2}.

\revnote{I fail to see the point here - the logic of the paper is: % {{{
1."open quantum systems can be modeled through a class of channels representing
he various types of decoherence. 2. Quantum circuits are a convenient
gate-based formalism that is well understood and has an extensive body of work,
and experimentally accessible.  3. We map (generalised) Pauli channels onto
quantum circuits, as a way to simulate them. \\
What I am missing is what do we learn by doing it which can't be learned by
actually performing the Pauli channel onto a quantum system? For example,
applying well controlled sx, sy, sz operations and their coherent parametric
combination using a photonic polarisation or path encoding is experimentally
trivial. Why would I need to decompose (not even effectively for any channel)
the Pauli channels to match a gate based description? Which class of problems
can we simulate that cannot be simulated otherwise? Is there any advantage (for
example in speed up, resources like ancillas required, resistance to unwanted
noise etc..?)\\
Intuitively, it seems incredibly harder resource wise to create the sum
$b|gamma\rangle$ superposition state needed for the ancilla, just for the sake
of having it into a "quantum circuit" perspective.} % }}}

@@

We can also see how an arbitrary Pauli channel acts on an arbitrary  
density matrix.
To do it, we substitute \eref{eq: Density Matrix} 
in \eref{eq: Pauli channel 1 qbit}:
\begin{equation}
\mathcal{E}(\rho) = \dfrac{1}{2}\sum_{\gamma,\alpha=0}^3 k_{\gamma} r_{\alpha} \sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma}.
\label{eq:Edepol}
\end{equation}
This can be simplified noting that 
\begin{equation}
\sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma} = A_{\alpha,\gamma} \sigma_{\alpha}, \;\;\;\;\; \text{with} \; A = \begin{pmatrix}
1 & 1 & 1 & 1\\
1 & 1 & -1 &-1 \\
1 & -1 & 1 & -1 \\
1 & -1 & -1 & 1
\end{pmatrix},
\end{equation}
which was obtained following the 
commutation relation of Pauli matrices $[\sigma_\alpha,\sigma_\beta] = 2 i
\epsilon_{\alpha \beta \gamma} \sigma_\gamma$ and $\epsilon_{\alpha \beta \gamma} \sigma_\gamma$ is 
the Levi-Civita tensor. Then, \eref{eq:Edepol} takes the form
% 
% 
% 
% 
% This can be simplified by using the following property of Pauli matrices:
% \revnote{Really confusing notation, I suggest use the much more well known relation using the Levi-Civita tensor.}
% 
% 
% \tbnote{Siguiendo la relaci\'on de conmutaci\'on, resulta que las Levi Civitas desaparecen y quedan s\'olo deltas de Dirac:}
% {\color{green}
% \begin{equation*}
% \sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma}  =   A_{\alpha,\gamma} \sigma_{\alpha}, \;\;\;\;\;\; \text{with} \; A_{\alpha, \gamma} = (2\delta_{ij}-1)(2\delta_{i0}-1)(2\delta_{j0}-1)
% \end{equation*}
% }
% 
% 
% \begin{equation}
% \label{eq: propiedad-pauli}
% \sigma_{\gamma} \sigma_{\alpha} \sigma_{\gamma} = A_{\alpha,\gamma} \sigma_{\alpha}, \;\;\;\;\; \text{with} \; A = \begin{pmatrix}
% 1 & 1 & 1 & 1\\
% 1 & 1 & -1 &-1 \\
% 1 & -1 & 1 & -1 \\
% 1 & -1 & -1 & 1
% \end{pmatrix},
% \end{equation}
% which leads to
\begin{equation}
\label{eq: rho-transformada}
\mathcal{E}(\rho) = \dfrac{1}{2} \sum_{\alpha} \left(\sum_{\gamma} A_{\alpha, \gamma} k_{\gamma} \right) r_{\alpha} \sigma_{\alpha}.
\end{equation}
\eref{eq: rho-transformada} once again has the form of \eref{eq: Density Matrix} \revnote{eq 2} \tbnote{No, sí es la eq 1}
but with components $\left( \sum_{\gamma} A_{\alpha,\gamma} k_{\gamma} \right) r_{\alpha}$.
This gives us another way of understanding Pauli channels
as operations that take each component $r_{\alpha}$
of the density matrix and multiplies them
by $\sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}$, 
that is:
\begin{equation}
\label{eq: multipliers}
r_{\alpha}  \xrightarrow[\text{Channel}]{\text{Pauli}}  \tau_{\alpha} r_{\alpha} ,\;\; \tau_{\alpha} := \sum_{\gamma} A_{\alpha,\gamma} k_{\gamma}.
\end{equation}
Notice that $\tau_0 = 1$, which is a consequence of $\sum_{\gamma}k_{\gamma}=1$
and ensures that after the channel, the resulting density matrix still has trace one. 
Furthermore, reverting the definition of $\tau_{\alpha}$ by using that $A^{-1} = \dfrac{1}{4} A$, 
we get that $k_{\gamma} = \dfrac{1}{4} \sum_{\alpha} A_{\alpha,\gamma} \tau_{\alpha}$.
Then, using that $k_{\gamma} \geq 0$
we get the following conditions on the multipliers $\tau_{\alpha}$:
\begin{eqnarray}
\label{eq: conditions-tetrahedron}
&1+\tau_i -\tau_j - \tau_k \geq 0,  \;\;\; \text{for $i,j,k$ different numbers in $\{1,2,3\}$}, \\
&1+\tau_1 + \tau_2 + \tau_3 \geq 0.
\end{eqnarray}

@@

\revnote{At this point I am missing a proper definition of a "quantum circuit"
consistent with the mathematical formalism. The authors start from the very
basic regarding Pauli channels, but give for granted the quantum circuit
formalism.}

In this section we propose a quantum circuit that simulates $N$-qubit Pauli
channels. 
A quantum circuit is a model for representing computation on a quantum system. 
This computation may include preparing the initial state of the system, applying unitary operations and measuring the whole system or parts of it~\cite{Rieffel}.
Our systems always consists of some number of qubits, that can be divided into
the ones to which we will apply the quantum channel, or ``main qubits''
and the ones that help us complete said operation, or ``ancilla qubits''.
The circuit implements a unitary operator $U$ on the system consisting of all
the qubits and then the effect on the main qubits is measured. 
Moreover, we implement the circuit for $N=1$ on a quantum computer
and analyze the results using the diamond norm.
We find that close to the depolarizing channel, the general circuit simulator 
can implement such channels with the highest fidelity.

@@

% {\color{green}
% Finally, we move to the main goal of this article, which is 
% simulating Pauli dynamical maps.
% 
Finally, we consider the simulation of 
Pauli dynamical maps.
% 
This is pretty much already solved, since
just as Pauli channels, Pauli dynamical maps can be implemented using the circuit
of \fref{Fig1}. However, there is one difference: 
the state to be created on the ancilla qubits now 
depends on a parameter $p$, and it is represented by the expression:
\begin{eqnarray}
\label{eq: parametrized state}
\sum_{\vec{\gamma}} b_{\vec{\gamma}}(p) |\vec{\gamma}\rangle.
\end{eqnarray}
Thus, we temporarily shift our focus from Pauli
channels and dynamical maps to the general problem of 
creating a circuit to generate a curve of
states like the one described in \eref{eq: parametrized state}.
